---
title: "Project 1"
author: "Koko Hall"
date: "2024-09-08"
output: html_document
---

```{r}
#Install packages 
library("tidyverse")

if (!require("tidytuesdayR", quietly = TRUE)) {
    install.packages("tidytuesdayR")
}

## Install the remotes package if you don't have it
if (!require("remotes", quietly = TRUE)) {
    install.packages("remotes")
}
## Install the tidytuesdayR package (from GitHub) if you don't have it
if (!require("tidytuesdayR", quietly = TRUE)) {
    remotes::install_github("dslc-io/tidytuesdayR")
}
```

```{r}
#download the data
tuesdata<- tidytuesdayR::tt_load("2022-01-18")
chocolate <- tuesdata$chocolate
```

```{r}
library("here")
library("tidyverse")

# tests if a directory named "data" exists locally
if (!dir.exists(here("data"))) {
    dir.create(here("data"))
}

# saves data only once (not each time you knit a R Markdown)
if (!file.exists(here("data", "chocolate.RDS"))) {
    url_csv <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-18/chocolate.csv"
    chocolate <- readr::read_csv(url_csv)

    # save the file to RDS objects
    saveRDS(chocolate, file = here("data", "chocolate.RDS"))
}

chocolate<- readRDS(here("data", "chocolate.RDS"))
as_tibble(chocolate)
glimpse(chocolate)
```

<span style="font-size: 20px;">Part 1 1. Make a histogram of the rating scores to visualize the overall distribution of scores. Change the number of bins from default to 10, 15, 20, and 25. Pick on the one that you think looks the best. Explain what the difference is when you change the number of bins and explain why you picked the one you did.

```{r}
ggplot(chocolate, aes(x= rating))+
  geom_histogram(bins=25)+
  ggtitle("Distribution scores of chocolate rating")
#With 25 bins, the data is divided into smaller intervals, each bin covers a smaller set of values. With 10 bins, the data is divided into larger intervals, covering a larger range of values. I chose an interval of 25, because you are able to distinguish rating scores better. I feel that this gives more detail about the distribution of chocolate ratings. 
```
<span style="font-size: 20px;">2. Consider the countries where the beans originated from. How many reviews come from each country of bean origin?

```{r}
bean_origin<- chocolate %>%
  group_by(country_of_bean_origin) %>%
  summarise(review_count= n())
bean_origin
```

<span style="font-size: 20px;">3.What is average rating scores from reviews of chocolate bars that have Ecuador as country_of_bean_origin in this dataset? For this same set of reviews, also calculate (1) the total number of reviews and (2) the standard deviation of the rating scores. Your answer should be a new data frame with these three summary statistics in three columns. Label the name of these columns mean, sd, and total.

```{r}
ecuador_summary<- chocolate %>%
  filter(country_of_bean_origin == "Ecuador") %>%
  summarise(
    mean=mean(rating),
    sd= sd(rating),
    total= n()
  )
ecuador_summary

```
<span style="font-size: 20px;">4.Which company location makes the best chocolate (or has the highest ratings on average) with beans from Ecuador?
```{r}
best_company <- chocolate %>%
  filter(country_of_bean_origin=='Ecuador') %>%
  group_by(company_location) %>%
  summarise(mean_rating= mean(rating)) %>%
  arrange(desc(mean_rating))
best_company
#Australia makes the best chocolate with beans from Ecuador 
```
<span style="font-size: 20px;">5. Calculate the average rating across all country of origins for beans. Which top 3 countries (for bean origin) have the highest ratings on average?

```{r}
best_origin <- chocolate %>%
  group_by(country_of_bean_origin) %>%
  summarise(mean_rating= mean(rating)) %>%
  arrange(desc(mean_rating))
best_origin
#Top 3 bean origins (by country): Tobago, China, Sao Tome & Principe
```
<span style="font-size: 20px;"> 6.Following up on the previous problem, now remove any countries of bean origins that have less than 10 chocolate bar reviews. Now, which top 3 countries have the highest ratings on average?
```{r}
Top3_rm10<- chocolate %>%
  group_by(country_of_bean_origin) %>%
  summarise(
    mean_rating= mean(rating),
    total_reviews=n() 
  )%>%
  filter(total_reviews>=10) %>%
  arrange(desc(mean_rating))
Top3_rm10
#Excluding those that have less than 10 chocolate revies, the top 3 countries are now Solomon Islands, Congo, and Cuba 
```

<span style="font-size: 20px;">7. For this last part, let’s explore the relationship between percent chocolate and ratings.
Use the functions in dplyr, tidyr, and lubridate to perform the following steps to the chocolate dataset:
1. Identify the countries of bean origin with at least 50 reviews. Remove reviews from countries are not in this list.
2. Using the variable describing the chocolate percentage for each review, create a new column that groups chocolate percentages into one of four groups: (i) <60%, (ii) >=60 to <70%, (iii) >=70 to <90%, and (iii) >=90% (Hint check out the substr() function in base R and the case_when() function from dplyr – see example below).
3.Using the new column described in #2, re-order the factor levels (if needed) to be starting with the smallest percentage group and increasing to the largest percentage group (Hint check out the fct_relevel() function from forcats).
4.For each country, make a set of four side-by-side boxplots plotting the groups on the x-axis and the ratings on the y-axis. These plots should be faceted by country.
```{r}
fiftyrevs<- chocolate %>%
  group_by(country_of_bean_origin) %>%
  summarise(
    total_reviews=n()
  )%>%
  filter(total_reviews>=50) 
fiftyrevs

chocolate_50 <- chocolate %>%
  filter(country_of_bean_origin %in% fiftyrevs$country_of_bean_origin)
```

```{r}
chocolate_50 <- chocolate_50 %>%
  mutate(
    cocoa_numeric = as.numeric(gsub("%", "", cocoa_percent)), 
    
    percent_group = case_when(
      cocoa_numeric < 60 ~ "<60%", 
      cocoa_numeric >= 60 & cocoa_numeric < 70 ~ ">=60 to <70%",
      cocoa_numeric >= 70 & cocoa_numeric < 90 ~ ">=70 to <90%",
      cocoa_numeric >= 90 ~ ">90%",
    ) %>%
    factor(levels = c("<60%", ">=60 to <70%",">=70 to <90%" , ">90%"))
  ) %>%
  arrange(percent_group)
chocolate_50
```



```{r}
ggplot(chocolate_50, aes(x = percent_group, y = rating)) +
  geom_boxplot() +  
  facet_wrap(~ country_of_bean_origin) +  
  labs(x = "Group", y = "Rating", title = "Boxplots of Ratings by Percent Group and Country of Bean Origin") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotates the labels to be diagonal
```

```{r}
chocolate_summary <- chocolate_50 %>%
  group_by(percent_group) %>%
  summarise(
    avg_rating = mean(rating, na.rm = TRUE),
    countries = paste(unique(country_of_bean_origin), collapse = ", ")
  ) %>%
  arrange(desc(avg_rating))

chocolate_summary


```

<span style="font-size: 20px;">On average, >=60 -70% cocoa percent was the most highly rated. The countries tend to agree with this finding. 




<span style="font-size: 20px;">Part 2. Join two datasets together
The goal of this part of the assignment is to join two datasets together. gapminder is a R package that contains an excerpt from the Gapminder data.

```{r}
library(gapminder)
head(gapminder)
```

```{r}
#Create a unique list of countries and their corresponding continent 
country_continent <- gapminder %>%
  select (country, continent) %>%
  distinct()
```


```{r}
#Join country_continent with chocolate
chocolate_continent <-chocolate %>%
  left_join(country_continent, by = c("country_of_bean_origin"="country")) %>%
  filter(country_of_bean_origin != "Blend")
chocolate_continent
```

```{r}
#Remove those with less than 10 reviews 
tenrevs<- chocolate_continent %>%
  group_by(country_of_bean_origin) %>%
  summarise(total_reviews = n()) %>%
  filter(total_reviews >= 10)

chocolate_continent_10 <- chocolate_continent %>%
  filter(country_of_bean_origin %in% tenrevs$country_of_bean_origin)

chocolate_continent_10
```


```{r}
#Fill in NAs
chocolate_continent_10 <- chocolate_continent_10 %>%
  mutate(continent = case_when(
    country_of_bean_origin == "Belize" ~ "Americas",
    country_of_bean_origin == "U.S.A." ~ "Americas",
    country_of_bean_origin == "Congo" ~ "Africa", 
    country_of_bean_origin == "Fiji" ~ "Oceania",
    country_of_bean_origin == "Grenada" ~ "Americas",
    country_of_bean_origin == "Papua New Guinea" ~ "Oceania",
    country_of_bean_origin == "Sao Tome" ~ "Africa",
    country_of_bean_origin == "Solomon Islands" ~ "Oceania",
    country_of_bean_origin == "St. Lucia"~ "Americas",
    country_of_bean_origin == "Trinidad" ~ "Americas",
    country_of_bean_origin == "Vanuatu" ~ "Oceania",
    TRUE ~ continent
  ))

chocolate_continent_10
```

```{r}
#Create violin plots 
ggplot(chocolate_continent_10, aes(x = continent, y = rating)) +
  geom_violin(aes(fill = continent), color = "black") +  
  geom_boxplot(width = 0.1, color = "black", alpha = 0.7, position = position_dodge(0.9)) +  # Box plot inside the violin
  scale_fill_manual(values = c("Americas" = "lightblue", "Asia" = "lightgreen", "Oceania" = "lightcoral", "Africa"= "purple")) +  # Custom colors
  labs(title = "Violin Plot with Box Plot of Ratings by Continent", x = "Continent", y = "Rating") +
  theme_classic()
```

<span style="font-size: 20px;">Part 3. Convert wide into long data. Tasks
We are going to create a set of features for us to plot over time. Use the functions in dplyr and tidyr to perform the following steps to the chocolate dataset:1.Create a new set of columns titled beans, sugar, cocoa_butter, vanilla, letchin, and salt that contain a 1 or 0 representing whether or not that review for the chocolate bar contained that ingredient (1) or not (0).2.Create a new set of columns titled char_cocoa, char_sweet, char_nutty, char_creamy, char_roasty, char_earthy that contain a 1 or 0 representing whether or not that the most memorable characteristic for the chocolate bar had that word (1) or not (0). For example, if the word “sweet” appears in the most_memorable_characteristics, then record a 1, otherwise a 0 for that review in the char_sweet column (Hint: check out str_detect() from the stringr package). 3.For each year (i.e. review_date), calculate the mean value in each new column you created across all reviews for that year. (Hint: If all has gone well thus far, you should have a dataset with 16 rows and 13 columns).4.Convert this wide dataset into a long dataset with a new feature and mean_score column.

```{r}

chocolate_ingred <- chocolate %>%
  mutate(
    beans = ifelse(grepl("B", ingredients), 1, 0),
    sugar = ifelse(grepl("S", ingredients), 1, 0),
    cocoa_butter = ifelse(grepl("C", ingredients), 1, 0),
    vanilla = ifelse(grepl("V", ingredients), 1, 0),
    letchin = ifelse(grepl("L", ingredients), 1, 0),
    salt = ifelse(grepl("Sa", ingredients), 1, 0)  
  )

```

```{r}
library(stringr)
chocolate_char <- chocolate_ingred %>%
  mutate(
    char_cocoa = ifelse(str_detect(most_memorable_characteristics, "cocoa"), 1, 0),
    char_sweet = ifelse(str_detect(most_memorable_characteristics, "sweet"), 1, 0),
    char_nutty = ifelse(str_detect(most_memorable_characteristics, "nutty"), 1, 0),
    char_creamy = ifelse(str_detect(most_memorable_characteristics, "creamy"), 1, 0),
    char_roasty = ifelse(str_detect(most_memorable_characteristics, "roasty"), 1, 0),
    char_earthy = ifelse(str_detect(most_memorable_characteristics, "earthy"), 1, 0)
  )

```

```{r}
#sort by year, mean value of scores 
mean_values_year <- chocolate_char %>%
  group_by(review_date) %>%
  summarise(
    mean_beans = mean(beans),
    mean_sugar = mean(sugar),
    mean_cocoa_butter = mean(cocoa_butter),
    mean_vanilla = mean(vanilla),
    mean_letchin = mean(letchin),
    mean_salt = mean(salt),
    mean_char_cocoa = mean(char_cocoa),
    mean_char_sweet = mean(char_sweet),
    mean_char_nutty = mean(char_nutty),
    mean_char_creamy = mean(char_creamy),
    mean_char_roasty = mean(char_roasty),
    mean_char_earthy = mean(char_earthy)
  )
mean_values_year
```

```{r}
#convert wide into long
chocolate_long <- mean_values_year %>%
  pivot_longer(
    cols = mean_beans:mean_char_earthy, 
    names_to = "feature",  # New column for feature
    values_to = "mean_score"  # New column for the mean scores
  )
chocolate_long
```

<span style="font-size: 20px;">Part 4. In this part of the project, we will continue to work with our now tidy song dataset from the previous part.Tasks:Use the functions in ggplot2 package to make a scatter plot of the mean_scores (y-axis) over time (x-axis). There should be one plot (facet) for each feature. For full credit, your plot should include:An overall title for the plot and a subtitle summarizing key trends that you found. Also include a caption in the figure with your name.Both the observed points for the mean_score, but also a smoothed non-linear pattern of the trend. All plots should be shown in the one figure. There should be an informative x-axis and y-axis label

```{r}

ggplot(chocolate_long, aes(x = review_date, y = mean_score)) +
  geom_point() +  # Scatter points for mean scores
  geom_smooth(method = "loess", se = FALSE, color= "purple") +  # Smoothed non-linear trend
  labs(
    title = "Mean Scores Over Time by Feature of Chocolate",
     subtitle = str_wrap(" Mean scores of ingredients: lecithin, vanilla, and cocoa butter decrease over time. Generally, mean scores of the most memorable characteristics, creamy, decreased over time, while sweet and cocoa have increased in the last five years.", width = 80),
    caption = "Koko Hall",
    x = "Review Date (Year)",  
    y = "Mean Feature Scores" 
  ) +
  facet_wrap(~feature, scales = "free_y") +  # Facet for each feature
  theme_light()


```

<span style="font-size: 20px;">Part 5.Make the worst plot you can!
This sounds a bit crazy I know, but I want this to try and be FUN! Instead of trying to make a “good” plot, I want you to explore your creative side and make a really awful data visualization in every way. :)Tasks: Using the chocolate dataset (or any of the modified versions you made throughout this assignment or anything else you wish you build upon it):Make the absolute worst plot that you can. You need to customize it in at least 7 ways to make it awful.In your document, write 1 - 2 sentences about each different customization you added (using bullets – i.e. there should be at least 7 bullet points each with 1-2 sentences), and how it could be useful for you when you want to make an awesome data visualization.

```{r}
ggplot(mean_values_year, aes(x = review_date, y = mean_char_creamy)) +
  geom_point()+ 
  geom_smooth(method = "loess", color= "white", se=FALSE) + 
  scale_y_continuous(limits= c(0, 10)) +
  scale_x_continuous(limits= c(1990, 2025)) + 
  labs (title= "Graph")+
  theme(
    axis.title.x  = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text (size=8),
    panel.background = element_rect(fill = "white"), # Light panel background
  )+
  annotate("text", x=2008, y=5, label= "Koko", size= 60, color = "red", fontface= "bold")

```

<span style="font-size: 20px;">1.The title is extremely small and not readable
2.The title is not informative. Titles help the viewer undrestand what
the graph is about.
3.There are no y or x axis labels 
4.The plot has a smoothed, non-linear line, but it is in white. It's very hard to see the line. 
5. The x-axis contains years that were before the study was conducted. There is a lot of unnecessary space.
6. The Y-axis is not scaled correctly, you aren't able to observe the trend in the data 
7. The caption is centerd with the creator's name and is blocking the contents of the graph



<span style="font-size: 20px;">Part 6. Make my plot a better plot!
The goal is to take my sad looking plot and make it better! If you’d like an example, here is a tweet I came across of someone who gave a talk about how to zhoosh up your ggplots.
```{r}
chocolate %>%
  ggplot(aes(
    x = as.factor(review_date),
    y = rating,
    fill = as.factor(review_date)  # Map fill to review_date
  )) +
  geom_violin(adjust = 1.5) +                            # Adjusted violin density for smoother plots
  scale_fill_manual(values = rainbow(length(unique(chocolate$review_date)))) +  # Use rainbow palette
  geom_boxplot(width = 0.1, color = "black", alpha = 0.5) +  # Overlay boxplots
  scale_x_discrete(breaks = seq(min(chocolate$review_date), max(chocolate$review_date), by = 5)) +    
  labs(
    title = "Chocolate Rating by Year",
    subtitle = str_wrap ("The range of chocolate ratings has narrowed over time. Medians scores have remained relatively stable ",  width = 80) ,
    x = "Review Year",
    y = "Rating",
    fill = "Review Year"
  ) +
  theme_minimal() +                                  # Apply a clean minimal theme
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better readability
```

<span style="font-size: 20px;">1. I added a clear title to tell the reader what the graph was about
2. The x and y axes needed to be re-labelled 
3. The years on the x axis were too crammed. I labelled the axis with less # of years
4. I tilted the x-axis years so they were more readable on the graph 
5. The monochromatic blue was hard to distinguish the different years. Different colors per year were easier to follow. 
6. I added a subtitle to describe the graph
7. I added boxplots to overlay the violin plots. We can measure the median and IQR values easier.


```{r}
options(width=120)
sessioninfo::session_info()

```
